---
title: "Holsenbeck_S_7"
author: "Stephen Synchronicity"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
always_allow_html: yes
output: 
  html_document: 
    css: C:\Users\Stephen\Documents\R\win-library\3.4\rmarkdown\rmarkdown\templates\neu_hwk\resources\styles.css
    highlight: zenburn
    keep_md: yes
    theme: readable
    toc_float: true
    toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,cache=TRUE, fig.align='center', fig.height=3.5, fig.width=5, tidy=TRUE, tidy.opts=list(width.cutoff=80))
rmarkdown::html_dependency_jquery()
rmarkdown::html_dependency_bootstrap("readable")
rmarkdown::html_dependency_jqueryui()
set.seed(1)
```
```{r Libraries, echo=FALSE, results='hide'}
library("tidyverse")
library("dplyr")
library("magrittr")
library("htmltools")
library("DT")
library("ggplot2")
```


<button data-toggle="collapse" data-target="#demo" class="btn">Homework Outline</button>
<div id="demo" class="collapse">
Is there an effect of Age on IQ? Please perform all calculations by hand using the equations in the lessons
unless otherwise specified.
1. Plot these four points using R.
2. Calculate the covariance between age and IQ.
3. Calculate their correlation. What does the number you get indicate?
4. Calculate the regression coefficients $\beta_0$ and $\beta_1$ and write out the equation of the best-fit line relating
age and IQ.
5. Calculate the predicted $\hat{y_i}$ for each $x_i$.
6. Calculate R$\beta_2$ from the TSS/SSE equation. How does it relate to the correlation? What does the
number you get indicate?
7. Calculate the standard error of $\beta_1$, and use that to test (using the t test) whether $\beta_1$ is significant.
8. Calculate the p-value for $\beta_1$ and interpret it.
9. Calculate the$\beta_9$5% CI for $\beta_1$ and interpret it.
$\beta_0$. Confirm your results by regressing IQ on Age using R.
$\beta_1$. Plot your points again using R, including the linear fit line with its standard error.
12. What are you final conclusions about the relationship between age and IQ?
1
</div>
#Homework 7

```{r 'Create Table'}
ageIQ <- tibble::tribble(
    ~Age,~IQ,
   23L,100L,
   18L,105L,
    10L,95L,
   45L,120L
  )
ageIQ.s  <- cbind(ageIQ,Sums=rowSums(ageIQ))
ageIQ.s <- rbind(ageIQ.s,Sums=colSums(ageIQ.s))
summary(ageIQ)
```

##1
<div class="q">Plot these four points using R.</div>
```{r '1'}
ggplot(data = ageIQ,mapping=aes(x=Age,y=IQ))+
geom_point()+
geom_smooth(method=lm)+
  ggtitle("Age V IQ")+
  theme(plot.title = element_text(hjust = 0.5))
```

##2
<div class="q">Calculate the covariance between age and IQ.</div>
<div class="a">A)
$\textrm{Cov}(x,y) = \frac{1}{(n-1) } \sum_{i} (x_{i} - \bar{x})(y_{i} - \bar{y})$  
$\textrm{Cov}(x,y) = \frac{1}{(4-1) } \sum_{i} [(23 - 24)(100 - 105)+(18 - 24)(105 - 105)+...+(45 - 24)(120 - 105)]$  
$\textrm{Cov}(x,y) = 153.\bar3$  
</div>
```{r '2'}
ageIQ.cov <- ageIQ %>% 
  mutate(S=(Age-mean(ageIQ$Age))*(IQ-mean(ageIQ$IQ))) %>% colSums()
ageIQ.cov[3]/3
```

