---
title: "Holsenbeck_S_8_9"
author: "Stephen Synchronicity"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
always_allow_html: yes
header-includes:
   - \usepackage{dcolumn}
output: 
  html_document: 
    css: C:\Users\Stephen\Documents\R\win-library\3.4\rmarkdown\rmarkdown\templates\neu_hwk\resources\styles.css
    highlight: zenburn
    keep_md: yes
    theme: readable
    toc_float: true
    toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,cache=TRUE, fig.align='center', fig.height=3.5, fig.width=5, tidy=TRUE, tidy.opts=list(width.cutoff=80))
rmarkdown::html_dependency_jquery()
rmarkdown::html_dependency_bootstrap("readable")
rmarkdown::html_dependency_jqueryui()
set.seed(1)
```
```{r Libraries, echo=FALSE, results='hide'}
library("tidyverse")
library("dplyr")
library("htmltools")
library("stringr")
library("forcats")
library("gridExtra")
library("magrittr")
```


<button data-toggle="collapse" data-target="#demo" class="btn">Homework Outline</button>
<div id="demo" class="collapse">
1	Describe your substantive interest and the general questions(s) you would like to answer (eg, “Does more education cause people to become more liberal?”). Be sure to frame it in a such a way that you are proposing a hypothesis (or multiple hypotheses) that might be either confirmed or disproven by the results of your analysis.
2	Describe the data set you have found, including its source, its contents, and why it was collected originally.
3	What is your dependent variable? Why are you interested in explaining it? What do you hypothesize are the major factors that influence or cause it?
4	What are your independent variables, and why have you chosen these? Prior to running your regression, what effects do you expect them to have on the dependent variable? Which of these variables do you think a˙ect other of the independent variables, and how might that a˙ect your final results?
5	Explain and show in detail how you rename and recode the variables you are examining, and what units each are measured in.
6	Before running a multiple regression, run a few bivariate regressions of Y on some of your X variables. What do you infer? Which of these do you think might change with the addition of multiple variables?
7	Run your full multiple regression using lm() and present your results using the output from the stargazer R package. Interpret the coefficients. What do they tell you substantively? Which variables seem to have the biggest substantive impact? Which ones could you actually change with some intervention, and how big a di˙erence do you think that could make?
8	How have any of the coefficients changed from the bivariate regressions? What can you infer from that?How do you think your various independent variables interact and a˙ect each other? Try to find an example where a variable appears signficant in the bivariate regression, but not in the full regression. Is this an example of a spurious or a chained causal pathway?
9	How does what you see match, or not, your hypotheses from (4)? Why did/didn’t it match what you expected?
10	What do the R2 and adjusted R2 tell you about your model?
11	How would you use one of the variable selection methods to choose a model with fewer variables? Select one of the methods (either one of the stepwise or criterion-based methods) and show which variables it would lead you to keep. Do you agree with its results?
12	What are your overall conclusions? What are the weaknesses of your results, and how could you improve them with better or different data?
13	Calculations (using R):
a.	Derive the coeÿcients from your regression using the (X0X)−1X0Y formula. (If you run into problems using solve(), try using ginv() instead, which does the same thing but is a bit more robust.)
b.	For one of the coeÿcients, confirm its p value as shown in the regression output using the coeÿcient, its standard error, and pt() in R.
c.	Calculate the R2 and adjusted R2 using R, and confirm that your results match the regression output.
d.	Calulate the F statistic using R and confirm it against the regression output.
14	Add at least one quadratic term into your model and interpret the results. Is it significant? What is the effect of a 1-unit increase in that variable at its mean value?
15	Add at least one interaction term to you model and interpret the results. Is it significant? What is the effect of a 1-unit increase in one of those interacted variables holding the other at its mean value?
16	Test either the model in 14 or the model in 15 using the F test for nested models. That is, estimate the full model with the variable and quadratic term, or the variable and interaction, and then estimate the reduced model without either, and run the F test to establish whether those variables significantly improve your model.
</div>


**Read Data Set 1**
```{r 'Read Data Set 1'}
US <- read_csv("C:\\Users\\Stephen\\Documents\\Northeastern\\PPUA 5301 - Introduction to Computational Statistics\\Homework\\Homework 8_9\\Nutrition__Physical_Activity__and_Obesity_-_Behavioral_Risk_Factor_Surveillance_System.csv")
# names(US)
(Qs <- unique(US$Question))
```

```{r 'Xform Dataset 1',eval=F}
US_Hlth <- US %>% filter(YearStart==2015 & Question %in% c("Percent of adults who engage in no leisure-time physical activity","Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)","Percent of adults aged 18 years and older who have an overweight classification","Percent of adults who report consuming vegetables less than one time daily") & is.na(Gender)==F & !LocationDesc %in% c("Guam","Puerto Rico","National")) %>%
  select(YearStart,LocationDesc, Question, Data_Value,Gender) %>% 
  group_by(LocationDesc, Question) %>%
  arrange(LocationDesc,Question) %>%
  summarize(Perc=mean(Data_Value))
colnames(US_Hlth) <- c("ST","Hlth_M","Perc")
US_Hlth$Hlth_M <- as.factor(US_Hlth$Hlth_M)
US_Hlth$Hlth_M <- US_Hlth$Hlth_M %>% fct_recode("0"="Percent of adults who engage in no leisure-time physical activity","150"="Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)","Veg"="Percent of adults who report consuming vegetables less than one time daily","Wgt"="Percent of adults aged 18 years and older who have an overweight classification")
US_Hlth<- US_Hlth %>% ungroup() %>%  filter(Hlth_M %in% c("0","150","Veg","Wgt")) %>% spread("Hlth_M","Perc")
```
```{r 'Xform Dataset 1 - Revised'}
US_Hlth <- US %>% filter(YearStart==2015 & is.na(Gender)==F & !LocationDesc %in% c("Guam","Puerto Rico","National")) %>%
  select(YearStart,LocationDesc, Question, Data_Value,Gender) %>% 
  group_by(LocationDesc, Question) %>%
  arrange(LocationDesc,Question) %>%
  summarize(Perc=mean(Data_Value))
colnames(US_Hlth) <- c("ST","Hlth_M","Perc")
US_Hlth$Hlth_M <- as.factor(US_Hlth$Hlth_M)
US_Hlth$Hlth_M <- US_Hlth$Hlth_M %>% fct_recode("wgtObs" = "Percent of adults aged 18 years and older who have obesity",
"wgtOvr" = "Percent of adults aged 18 years and older who have an overweight classification",
"conFru" = "Percent of adults who report consuming fruit less than one time daily",
"conVeg" = "Percent of adults who report consuming vegetables less than one time daily",
"actWL" = "Percent of adults who engage in muscle-strengthening activities on 2 or more days a week",
"act150" = "Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)",
"act150WL" = "Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic physical activity and engage in muscle-strengthening activities on 2 or more days a week",
"act300" = "Percent of adults who achieve at least 300 minutes a week of moderate-intensity aerobic physical activity or 150 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)",
"act0" = "Percent of adults who engage in no leisure-time physical activity")
US_Hlth<- US_Hlth %>% ungroup() %>%  spread("Hlth_M","Perc")
```

**Read Dataset 2**
```{r 'Read Dataset 2'}
Party <- tibble::tribble(
                            ~State, ~Republican.lean.Rep., ~No.lean, ~Democrat.lean.Dem., ~Sample.Size,
                         "Alabama",                 "52%",    "13%",               "35%",          511,
                          "Alaska",                 "39%",    "29%",               "32%",          310,
                         "Arizona",                 "40%",    "21%",               "39%",          653,
                        "Arkansas",                 "46%",    "16%",               "38%",          311,
                      "California",                 "30%",    "21%",               "49%",         3697,
                        "Colorado",                 "41%",    "17%",               "42%",          504,
                     "Connecticut",                 "32%",    "18%",               "50%",          377,
                        "Delaware",                 "29%",    "17%",               "55%",          301,
            "District of Columbia",                 "11%",    "15%",               "73%",          303,
                         "Florida",                 "37%",    "19%",               "44%",         2020,
                         "Georgia",                 "41%",    "18%",               "41%",          968,
                          "Hawaii",                 "28%",    "20%",               "51%",          312,
                           "Idaho",                 "49%",    "19%",               "32%",          320,
                        "Illinois",                 "33%",    "19%",               "48%",         1326,
                         "Indiana",                 "42%",    "20%",               "37%",          654,
                            "Iowa",                 "41%",    "19%",               "40%",          330,
                          "Kansas",                 "46%",    "23%",               "31%",          307,
                        "Kentucky",                 "44%",    "13%",               "43%",          439,
                       "Louisiana",                 "41%",    "16%",               "43%",          465,
                           "Maine",                 "36%",    "17%",               "47%",          303,
                        "Maryland",                 "31%",    "14%",               "55%",          644,
                   "Massachusetts",                 "27%",    "17%",               "56%",          704,
                        "Michigan",                 "34%",    "19%",               "47%",          982,
                       "Minnesota",                 "39%",    "15%",               "46%",          563,
                     "Mississippi",                 "44%",    "14%",               "42%",          309,
                        "Missouri",                 "41%",    "18%",               "42%",          642,
                         "Montana",                 "49%",    "21%",               "30%",          312,
                        "Nebraska",                 "47%",    "17%",               "36%",          312,
                          "Nevada",                 "37%",    "18%",               "46%",          314,
                   "New Hampshire",                 "35%",    "20%",               "44%",          303,
                      "New Jersey",                 "30%",    "19%",               "51%",          886,
                      "New Mexico",                 "37%",    "15%",               "48%",          312,
                        "New York",                 "28%",    "19%",               "53%",         1966,
                  "North Carolina",                 "41%",    "17%",               "43%",         1022,
                    "North Dakota",                 "50%",    "18%",               "33%",          338,
                            "Ohio",                 "42%",    "18%",               "40%",         1132,
                        "Oklahoma",                 "45%",    "15%",               "40%",          391,
                          "Oregon",                 "32%",    "21%",               "47%",          419,
                    "Pennsylvania",                 "39%",    "15%",               "46%",         1366,
                    "Rhode Island",                 "30%",    "22%",               "48%",          305,
                  "South Carolina",                 "43%",    "18%",               "39%",          495,
                    "South Dakota",                 "53%",    "10%",               "37%",          305,
                       "Tennessee",                 "48%",    "15%",               "36%",          661,
                           "Texas",                 "39%",    "21%",               "40%",         2535,
                            "Utah",                 "54%",    "16%",               "30%",          315,
                         "Vermont",                 "29%",    "14%",               "57%",          306,
                        "Virginia",                 "43%",    "18%",               "39%",          882,
                      "Washington",                 "33%",    "23%",               "44%",          714,
                   "West Virginia",                 "43%",    "16%",               "41%",          309,
                       "Wisconsin",                 "42%",    "16%",               "42%",          600,
                         "Wyoming",                 "57%",    "18%",               "25%",          316
           )
```
**Xform Dataset 2**
```{r 'Xform Dataset 2'}
Party <- lapply(Party, gsub, pattern='\\%', replacement='')
Party <- as.data.frame(Party,stringsAsFactors = F,)
reclass <- function(df, vec){
  df[] <- Map(function(x, f){
    #switch below shows the accepted values in the vector
    #you can modify it and/or add more
    f <- switch(f,
                as.is  = 'force',
                factor = 'as.factor',
                num    = 'as.numeric',
                char   = 'as.character')
    #takes the name of the function and fetches the function
    f <- get(f)
    #apply the function
    f(x)
  },
      df,
      vec)
df
}
Party <- reclass(Party, c('char','num','num','num','num'))
names(Party)
colnames(Party) <- c("ST","R","N","D","ns")
```

**Join Tables for Analysis**
```{r 'Join Tables for Analysis'}
Pty_Hlth <- left_join(Party,US_Hlth,by="ST")
rm(US_Hlth,Party)
```

**Prelim Analysis - Revised**
```{r 'Prelim Analysis - Revised', fig.width=10,fig.height=5}
Pty_Hlth<- Pty_Hlth %>% arrange(R)
ggact <- ggplot(data=Pty_Hlth,mapping=aes(x=reorder(ST,R),y=`R`))+
  geom_point()+
  geom_point(mapping=aes(y=`actWL`),color="red",size=2)+
  geom_point(mapping=aes(y=`act150`),color="orange",size=2)+
  geom_point(mapping=aes(y=`act150WL`),color="#C42B2D",size=2)+
  geom_point(mapping=aes(y=`act300`),color="blue",size=2)+
  geom_point(mapping=aes(y=`act0`),color="black",size=2)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1,lineheight = 1))
ggwgt <- ggplot(data=Pty_Hlth,mapping=aes(x=reorder(ST,R),y=`R`))+
  geom_point()+
  geom_point(mapping=aes(y=`wgtObs`),color="red",size=2)+
  geom_point(mapping=aes(y=`wgtOvr`),color="#A32D2D",size=2)+
  theme(legend.position="none",axis.text.x = element_text(angle = 90, hjust = 1,lineheight = 1))
ggcon <- ggplot(data=Pty_Hlth,mapping=aes(x=reorder(ST,R),y=`R`))+
  geom_point()+
  geom_point(mapping=aes(y=`conFru`),color="red",size=2)+
  geom_point(mapping=aes(y=`conVeg`),color="#B88037",size=2)+
  theme(legend.position="none",axis.text.x = element_text(angle = 90, hjust = 1,lineheight = 1))
#The Wgt variable is obviously not correlated, so this will be left out of further analyses
ggact
ggwgt
ggcon
```

## 1
<div class="q">	Describe your substantive interest and the general questions(s) you would like to answer (eg, “Does more education cause people to become more liberal?”). Be sure to frame it in a such a way that you are proposing a hypothesis (or multiple hypotheses) that might be either confirmed or disproven by the results of your analysis.
</div>
<div class="a">
Selecting the data for comparison was quite difficult given the restraints of having the data not include temporal or categorical data. All of my interests included data that had a temporal index or categorical data involved. I wanted to look at health related measures, and decided to carry on the question from the class theme about partisanship, hence the topic of the two datasets. The data used herein fits the restraints and *might* prove to yield interesting correlations.
<br>
The hypothesis is a question of whether there is a correlation between various health behaviors and party affiliation within states.
<blockquote>$H_1$ Is the proportion of individuals whom qualify as ?
$H_2$ Is the proportion of individuals meeting minimum recommendations for physical activity correlated with a party affilation?
$H_3$ Is the proportion of individuals exercising well beyond minimum recommendations for physical activity correlated with a party affilation</blockquote>
*After Revision*
$H_1$ Is the "Percent of adults aged 18 years and older who have obesity" correlated with identifyng as Republican? If so, what is the sign of $\beta_1$?
$H_2$ Is the "Percent of adults who report consuming fruit less than one time daily" correlated with a identifying as Republican? If so, what is the sign of $\beta_2$?
$H_3$ Is the "Percent of adults who achieve at least 300 minutes a week of moderate-intensity aerobic physical activity or 150 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)" a predictor variable for identifying as Republican? IF so what is the sign of $\beta_3$?
</div>
## 2
<div class="q">	Describe the data set you have found, including its source, its contents, and why it was collected originally.
</div>
<div class="a">
The data with proportion of the poulation living a lifestyle is a subset of the <a href="https://catalog.data.gov/dataset/nutrition-physical-activity-and-obesity-behavioral-risk-factor-surveillance-system" target="_blank">Nutrition, Physical Activity, and Obesity - Behavioral Risk Factor Surveillance System data set</a> from the CDC. The data is originally gathered to track behavioral risk factors across different demographic strata by state.
<br>
The data about party affilation by state came from the <a href="http://www.pewforum.org/religious-landscape-study/compare/party-affiliation/by/state/" target="_blank">Party affiliation by State (2014)</a> data set provided by the Pew Research Center. This particular data was originally collected as part of the <a href="http://www.pewforum.org/religious-landscape-study/" target="_blank">Religious Landscape Study</a>
</div>
## 3
<div class="q">	What is your dependent variable? Why are you interested in explaining it? What do you hypothesize are the major factors that influence or cause it?
</div>
<div class="a">The dependent variable is the percentage of the state pop. identfied as Republican, or Republican leaning. The variable is complex with numerous influencing factors. During the in-class explorations we have looked at potential influencing factors such as race, income, and education level on party affiliation via data from the ANES survey.
<br>
Intuitively, party affiliation could also be influenced or predicted in part by religion, affiliation of parents, place of residence, affiliation of peers in social network, or potentially to particular psychological traits such as openness.
<br>
In this study, we'll be looking at the percentage of each state's pop. exhibiting particular health behaviors, discussed further below. An individual presenting a risk factor of obesity could be due to lack of exposure to quality health related information while in the education system and/or lack of exposure to quality health related information sources. It could also be related to lack of access to health care services, medical professionals etc. If a person is living a lifestyle that involves behaviors identified as risk factors, there are often concurrent cognitive indicators resulting from poor health. 
<br>For example, there's evidence relating <a href="https://scholar.google.com/scholar?q=physical+exercise+and+executive+function&hl=en&as_sdt=0&as_vis=1&oi=scholart&sa=X&ved=0ahUKEwiDprWHwKvXAhWD5iYKHatJCmAQgQMIJDAA" target="_blank">physical exercise and executive function</a>, and <a href="https://scholar.google.com/scholar?hl=en&as_sdt=0%2C22&as_vis=1&q=physical+activity+on+cognition&oq=physical+exercise+and+executive+function" target="_blank">physical exercise to cognition</a>. Thus, there might be some explanatory value in relating activity levels (or lack thereof) to political party affiliation that would undoubtedly be controversial if the evidence exists. 
With regard to whether or not the "Percent of adults who report consuming fruit less than one time daily" influences the percent of the state pop. identifying as Republican it is likely that the two are not related. However, this variable was the only variable from the dietary behaviors questions that did not exhibit colinearity with variables from the other classes.
</div>
## 4
<div class="q">	What are your independent variables, and why have you chosen these? Prior to running your regression, what effects do you expect them to have on the dependent variable? Which of these variables do you think affect other of the independent variables, and how might that affect your final results?
</div>
<div class="a">
<blockquote>
This brief analysis will look at whether there is correlation between the following independent variables:
<ul style="list-style-type:none">
<li>$x_1$: % of pop. living sedentary lifestyle</li>
<li>$x_2$: % of pop. living lifestyle meeting minimum physical activity recommendations</li>
<li>$x_3$: % of pop. that consumes vegetables less than 1 time daily</li>
</ul>
and the dependent variable $y$: the identification as Republican.
<br>
There might be a positive correlation between the percent of a state's pop. that exhibits health behaviors identified as risk factors $x_1,x_3$ and the percent of a state's pop. that identifies as Republican or Republican leaning. There is likely an inverse colinear relationship between $x_1$ and $x_2$ because they are mutually exclusive categories along a scale of 5 categories (not all of which are mutually exclusive.) See variable Qs, for the original categories. There could also be a correlation between states with a larger percentage of the population meeting minimum physical activity recommendations, and those whom are consuming vegetables more than 1 time a day, due to a theoretical common variable of health knowledge.</blockquote>
*After Revision*
This brief analysis will look at whether there is correlation between the following independent variables:
<ul style="list-style-type:none">
<li>$x_1$: % of adults aged 18 years and older who have obesity</li>
<li>$x_2$: % of adults who report consuming fruit less than one time daily</li>
<li>$x_3$: % of adults who achieve at least 300 minutes a week of moderate-intensity aerobic physical activity or 150 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)</li>
</ul>
and the dependent variable $y$: the percent of the state pop. identifying as Republican.
The % of adults aged 18 years and older who have obesity ($x_1$) will likely correlate positively with the % of the state pop. identified as Republican. The % of adults who report consuming fruit less than one time daily will likely have no correlation with % of the state pop. identified as Republican. There will likely be an negative correlation between % of adults who achieve at least 300 minutes a week of moderate-intensity aerobic physical activity or 150 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination). However, this variable is likely more related to the latitude, and longitude of a state (as a predictor of outdoor weather patterns), as well as the popularity and diversity of various athletic recreational activities within the state.
</div>
## 5
<div class="q">	Explain and show in detail how you rename and recode the variables you are examining, and what units each are measured in.
</div>
<div class="a">
The variables have been recoded in the chunks entitled 
<div style="text-decoration: line-through">Xform Dataset 1"</div> Xform Dataset 1 - Revised & "Xform Dataset 2." The renaming and recoding procedures are explained below:
*Dataset 1*
The data is filtered for the year 2015, the only year with data for each question that is closest to the year in which the data was gathered in dataset 2, the relevant questions, the 50 states plus the District of Columbia (as in dataset 2), and the gender stratification. The "Questions" column from the original dataset is renamed Hlth_M (Health Measures). The percentages of the pop. reporting a yes to the question are filtered by the gender stratification, such that only 2 percentages need be averaged to get a single % for the entire population. The answers to the questions are then recoded using the forcats package, see the chunk entitled Xform Dataset 1 - Revised for details on variable names. The factors are then spread into columns.
*Dataset 2*
The dataset variables were transformed as follows:
<ol>
<li>R="Republican/Republican Leaning"</li>
<li>D="Democrat/Democrat Leaning"</li>
<li>N="No affiliation"</li>
</ol>
The variables were reclassed to characters or numbers where necessary. The tables are joined by State.
</div>

## 6
<div class="q">	Before running a multiple regression, run a few bivariate regressions of Y on some of your X variables. What do you infer? Which of these do you think might change with the addition of multiple variables?
</div>
<div class="a">
<blockquote>
*These linear regressions without results were used to identify that the variables in question were colinear. The assignment as then revised. The revised answer to this question begins where the blockquote ends.*
```{r '2-Var Regressions',eval=F}
`r0`<- summary(lm(`R`~`0`,Pty_Hlth))
`r150`<- summary(lm(`R`~`150`,Pty_Hlth))
`rv`<- summary(lm(`R`~`Veg`,Pty_Hlth))
`150v0` <- summary(lm(`0`~`150`,Pty_Hlth))
`150v0`
```
<p>Before running a multiple regression, run a few bivariate regressions of Y on some of your X variables. What do you infer? Which of these do you think might change with the addition of multiple variables?
<br>
While there is statistical significance at the 95% confidence level in the correlations between R and 150, and Veg, it can also be noted the $R^2$ values indicate that little of the variation is explained by the model. As variables are added in a multiple variable regression, the $R^2$ value will inevitably increase, though it is important to note the low $R^2$ in these 2-Variable models.
<br>
The comparison of `0` and `150` show that the variables are indeed colinear, and will be confounding variables if combined. It may be necessary to add an additional health variable that is not colinear with any of the other variables.</p>
```{r 'Add Statistic on Obesity for Multiple Linear Regression',eval=F}
USob <- US %>% filter(YearStart==2015 & Question %in% c("Percent of adults aged 18 years and older who have obesity") & is.na(Gender)==F & !LocationDesc %in% c("Guam","Puerto Rico","National")) %>%
  select(YearStart,LocationDesc, Question, Data_Value,Gender) %>% 
  group_by(LocationDesc, Question) %>%
  arrange(LocationDesc,Question) %>%
  summarize(Perc=mean(Data_Value))
USob <- USob[,-2]
names(USob) <- c("ST","Ob")
Pty_Hlth <- left_join(Pty_Hlth,USob,by="ST")
ROb <- summary(lm(`R`~`Ob`,Pty_Hlth))
ROb
rm(USob)
`0Ob` <- summary(lm(`0`~`Ob`,Pty_Hlth))
`0Ob`
```

It appears that obesity is also colinear with a sedentary lifestyle, logically, this linkage makes sense. Let's have a look at the fruit variable.
```{r 'Add Fruit Variable',eval=F}
USF <- US %>% filter(YearStart==2015 & Question %in% c("Percent of adults who report consuming fruit less than one time daily") & is.na(Gender)==F & !LocationDesc %in% c("Guam","Puerto Rico","National")) %>%
  select(YearStart,LocationDesc, Question, Data_Value,Gender) %>% 
  group_by(LocationDesc, Question) %>%
  arrange(LocationDesc,Question) %>%
  summarize(Perc=mean(Data_Value))
USF <- USF[,-2]
names(USF) <- c("ST","Fruit")
Pty_Hlth <- left_join(Pty_Hlth,USF,by="ST")
(VF <- summary(lm(Veg~Fruit,Pty_Hlth)))
#Colinear
(F150 <- summary(lm(Fruit~`150`,Pty_Hlth)))
#Colinear
(FOb <- summary(lm(Fruit~Ob,Pty_Hlth)))
#Colinear```
```
Healthy lifestyle traits show high degrees of correlation, while risk factors are also highly correlated. 
```{r 'Add Weight Lifting Variable',eval=F}
USWl <- US %>% filter(YearStart==2015 & Question %in% c("Percent of adults who engage in muscle-strengthening activities on 2 or more days a week") & is.na(Gender)==F & !LocationDesc %in% c("Guam","Puerto Rico","National")) %>%
  select(YearStart,LocationDesc, Question, Data_Value,Gender) %>% 
  group_by(LocationDesc, Question) %>%
  arrange(LocationDesc,Question) %>%
  summarize(Perc=mean(Data_Value))
USWl <- USWl[,-2]
names(USWl) <- c("ST","Wl")
Pty_Hlth <- left_join(Pty_Hlth,USWl,by="ST")
(VFWl <- summary(lm(Veg~Wl,Pty_Hlth)))
#Colinear
(Wl150 <- summary(lm(Wl~`150`,Pty_Hlth)))
#Colinear
(WlOb <- summary(lm(Wl~Ob,Pty_Hlth)))
#Colinear
```

It appears that it will be necessary to start back at the beginning, extract all of the health related questions, spread them to individual variables and graph them. By graphing all of the variables it may allow them to be differentiated to avoid colinearity.</blockquote>
*After Revision*
After revision, the three variables that appear not to be colinear are:
<ol>
<li>"% of adults aged 18 years and older who have obesity", <u>wgtObs</u></li>
<li>"% of adults who report consuming fruit less than one time daily", <u>conFru</u></li>
<li>"% of adults who achieve at least 300 minutes a week of moderate-intensity aerobic physical activity or 150 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)", <u>act300</u></li>
</ol>
</div>
## 7
<div class="q">	Run your full multiple regression using lm() and present your results using the output from the stargazer R package. Interpret the coefficients. What do they tell you substantively? Which variables seem to have the biggest substantive impact? Which ones could you actually change with some intervention, and how big a difference do you think that could make?
</div>
<div class="a">

</div>
## 8
<div class="q">	How have any of the coefficients changed from the bivariate regressions? What can you infer from that? How do you think your various independent variables interact and affect each other? Try to find an example where a variable appears signficant in the bivariate regression, but not in the full regression. Is this an example of a spurious or a chained causal pathway?
</div>
<div class="a">
</div>
## 9
<div class="q">	How does what you see match, or not, your hypotheses from (4)? Why did/didn’t it match what you expected?</div>
<div class="a">
</div>
## 10
<div class="q">	What do the $R^2$ and adjusted $R^2$ tell you about your model?</div>
<div class="a">
</div>
## 11
<div class="q">	How would you use one of the variable selection methods to choose a model with fewer variables? Select one of the methods (either one of the stepwise or criterion-based methods) and show which variables it would lead you to keep. Do you agree with its results?
</div>
<div class="a">
</div>
## 12
<div class="q">	What are your overall conclusions? What are the weaknesses of your results, and how could you improve them with better or different data?
</div>
<div class="a">
</div>
## 13
<div class="q">	Calculations (using R):
a.	Derive the coefficients from your regression using the (X0X)−1X0Y formula. (If you run into problems using solve(), try using ginv() instead, which does the same thing but is a bit more robust.)
b.	For one of the coefficients, confirm its p value as shown in the regression output using the coeffcient, its standard error, and pt() in R.
c.	Calculate the R2 and adjusted R2 using R, and confirm that your results match the regression output.
d.	Calulate the F statistic using R and confirm it against the regression output.
</div>
<div class="a">
</div>
## 14
<div class="q">	Add at least one quadratic term into your model and interpret the results. Is it significant? What is the effect of a 1-unit increase in that variable at its mean value?</div>
<div class="a">
</div>
## 15
<div class="q">	Add at least one interaction term to you model and interpret the results. Is it significant? What is the effect of a 1-unit increase in one of those interacted variables holding the other at its mean value?</div>
<div class="a">
</div>
## 16
<div class="q">	Test either the model in 14 or the model in 15 using the F test for nested models. That is, estimate the full model with the variable and quadratic term, or the variable and interaction, and then estimate the reduced model without either, and run the F test to establish whether those variables significantly improve your model.</div>
<div class="a">
</div>

